Updated Constraints on Self-Interacting Dark Matter from Supernova 1987A

### 2017-01-05 draft.tex
Wrote a brief, new section 2 that serves to introduce our notation.

SIDM models in which large self-interaction cross sections are mediated by sufficiently light 
bosons ($M \lesssim 100$~GeV) can be constrained astrophysically using supernovae, particularly 
SN1987A. Light gauge bosons will be produced within the hot supernovae core and 
radiated from the supernova. This non-standard energy loss mechanism can result in an energy loss rate 
from the supernova core that is inconsistent with observations of SN1987A, analogous to the 
classic constraint on axions \cite{raffelt96_book}. This mechanism was already exploited by 
\cite{dent_etal12} (and \cite{rrapaj_reddy16}, with a slight modification) (\arz{Cameron, add others of which you may be aware.}) 
to constraint dark photon models of SIDM in which the self-interaction arises from a Yukawa potential and the 
vector boson is kinetically mixed with the standard model photon.

As we mentioned above, we find SN1987A constraints on dark photon SIDM models that are considerably different 
from those of \cite{dent_etal12}. If correct, this has important consequences. The Snowmass white paper by Kaplinghat, 
Tulin, and Yu \cite{kaplinghat_etal13_whitepaper} nicely summarizes a variety of constraints on dark photon models of 
SIDM. One point that is clear from Ref.~\cite{kaplinghat_etal13_whitepaper} is that there is only a narrow range of 
viable photon mixing parameters that can lead to production of SIDM in the early universe, while evading all constraints, 
including the constraint of \cite{dent_etal12}. Our results suggest that the SN1987A constraints from Ref.~\cite{dent_etal12} 
are too restrictive. The less restrictive constraints that we quote open up a space of unrestricted mixing parameters that 
ranges over two orders of magnitude. Interestingly, the parameter space opened by our updated constraints corresponds to a 
space that may a significant impact on cosmological structure formation, particularly the structures of dark matter halos and galaxies.

@@

@@

\begin{figure}[th]
\includegraphics[width=9cm]{stages.eps}
\caption{The changes in the bounds generated as each stage of the calculation departs from Ref.~\cite{dent_etal12}. Interestingly the kinematics have apparently relatively little effect.}
\end{figure}

@@

\begin{figure}[th]
	\includegraphics[width=9cm]{endtoend.png}
	\caption{A comparison of the final result for the excluded region with the result from Ref.~\cite{dent_etal12}. The excluded region is clearly significantly less restrictive, and opens up the region of parameter space for $\epsilon \approx  1e-8-9 $.}
\end{figure}

>>>>

SIDM models in which large self-interaction cross sections are mediated by sufficiently light 
bosons ($M \lesssim 100$~GeV) can be constrained astrophysically using supernovae, particularly 
SN1987A. Light gauge bosons will be produced within the hot supernovae core and 
radiated from the supernova. This non-standard energy loss mechanism can result in an energy loss rate 
from the supernova core that is inconsistent with observations of SN1987A, analogous to the 
classic constraint on axions \cite{raffelt96_book}. This mechanism was already exploited by 
\cite{dent_etal12} (and \cite{rrapaj_reddy16}, with a slight modification) (\arz{Cameron, add others of which you may be aware.}) 
to constraint dark electromagnetism models of SIDM in which the self-interaction arises from a Yukawa potential and the 
vector boson is kinetically mixed with the standard model photon.

As we mentioned above, we find SN1987A constraints on dark photon SIDM models that are considerably different 
from those of \cite{dent_etal12}. If correct, this has important consequences. The Snowmass white paper by Kaplinghat, 
Tulin, and Yu \cite{kaplinghat_etal13_whitepaper} nicely summarizes a variety of constraints on dark electromagnetism models of 
SIDM. One point that is clear from Ref.~\cite{kaplinghat_etal13_whitepaper} is that there is only a narrow range of 
viable photon mixing parameters that can lead to production of SIDM in the early universe, while evading all constraints, 
including the constraint of \cite{dent_etal12}. Our results suggest that the SN1987A constraints from Ref.~\cite{dent_etal12} 
are too restrictive. The less restrictive constraints that we quote open up a space of unrestricted mixing parameters that 
ranges over two orders of magnitude. Interestingly, the parameter space opened by our updated constraints corresponds to a 
space that may a significant impact on cosmological structure formation, particularly the structures of dark matter halos and galaxies.

@@

We consider constraints on SIDM specifically within the context of dark electromagnetism models. 
Dark electromagnetism models are models in which a hidden, dark, sector contains a broken U(1)$'$ 
symmetry and the U(1)$'$ gauge boson is kinetically mixed with the standard model photon. For the 
purposes of this study, this is important because it demands that the Lagrangian contains terms such as 
%
\begin{equation}
\mathcal{L}_{\mathrm{int}} = g_{\chi} \bar{\chi} \, \tilde{\slashed{A'} } \chi + q \bar{f}\, \tilde{\slashed{A} } f, 
\end{equation}
%
where $\chi$ is the dark matter, $g_{\chi}$ is the dark coupling, $\tilde{A}'$ is the dark gauge boson, 
$f$ is a standard model fermion of charge $q$, and $\tilde{A}$ is the standard model gauge boson. 
The kinetic mixing causes the $\tilde{A}$ to be an admixture of the massless photon, $A$, and 
the dark photon, $A'$ of mass $m_{\mathrm{A'}}$, so that the dark matter particles are coupled to the standard model 
fermions with a coupling constant $\epsilon q$, where $\epsilon$ is the kinetic mixing parameter. 
The first term in this interaction gives rise to the dark matter self-interactions.

Dark gauge bosons are produced in astrophysical environments such as supernova 
cores via brehmsstrahlung off of standard model particles. This brehmsstrahlung 
occurs through the $\epsilon q$ coupling. The rate of brehmsstrahlung 
also depends upon the mass of the $A'$. Consequently, 
supernova cooling can constrain the mixing $\epsilon$ as a function of 
$m_{A'}$ for such models and producing such a constraint is the 
primary aim of this paper.

@@

\begin{figure}[th]
\includegraphics[width=9cm]{stages.eps}
\caption{The changes in the bounds generated as each stage of the calculation departs from Ref.~\cite{dent_etal12}. Interestingly the kinematics have apparently relatively little effect.
\arz{Cameron, several to-do items for this figure. On the y-axis, the symbol should be "$\epsilon$" rather than 
"e". Can we write numbers as "$10^{-9}$" rather than "1e-09" and so on? Can we make the axis labels and 
tick mark labels larger? Can you make all of the lines thicker? Can you make each of the four lines a distinct color?}
}
\end{figure}

@@

\begin{figure}[th]
	\includegraphics[width=9cm]{endtoend.png}
	\caption{A comparison of the final result for the excluded region with the result from Ref.~\cite{dent_etal12}. The excluded region is clearly significantly less restrictive, and opens up the region of parameter space for $\epsilon \approx  1e-8-9 $.
\arz{Cameron, some similar comments as for Figure 1. Can we write numbers as "$10^{-9}$" rather 
than "1e-09" and so on? Can we make the axis labels and tick mark labels larger?}	
	}
\end{figure}
